# Conclusion

2020 has been the year of significant change in everyone's life forever to be written down as the year of COVID-19, but will there be more to come? Throughout this project we aimed to answer a few key questions and discovered some more interesting problems in the process. Below is a refresher of those questions and our conclusions from our results:  

## Initial Questions
**What are some major COVID-19 patterns (cases and deaths) in the US, in relation to major domestic events?**

In our analysis we noticed some alarming patterns and confirmed our suspicions that major holidays seemed to have an effect on the increasing the spread of COVID-19. The majority of states saw a rise of positive cases following Memorial Day weekend. In addition, we saw significant upticks following the protests against police brutality and Thanksgiving. However, this only paints one side of the story. Further analysis showed that positive cases were visibly correlated with the number of tests. This result, in fact, shows that the rise due to holidays seems to be just blips in the grand scheme of COVID-19 spread in the US. Moreover, it seems like we are hitting the exponential rise portion of the pandemics growth. As discussed earlier, the United States has had an upward trend until October, followed by an exponential increase similar to that of Illinois. Almost every state regardless of their performance in the summer months has exploded exponentially in the last two months. Below is the national trend with 7-day average:

```{r}
daily_us <- read_csv("daily_us.csv", col_types = cols(date = col_date(format = "%Y%m%d")))
daily_us$positiveIncrease_7 <- rollmean(daily_us$positiveIncrease, k=7, fill=NA)
p <- ggplot(data = daily_us, aes(date, positiveIncrease_7))
p + geom_line(alpha = 2, col = "blue") + theme_classic() + labs(y = "Daily Positive Cases", title = "COVID-19 Cases in the US are Increasing Exponentially")
```

**What are some major COVID-19 Patterns (cases and deaths) when stratified by states/regions, and if necessary, specific counties of the state?**

We stratified states based on the magnitude of tests and relative to their population and found that the "high" states, although usually performing worse in number of cases, did not trend that differently from "medium", and "low" states. This observation confirms the final result from our previous question that regardless of seasonality, holidays, or individual state performance, the US is heading towards the and exponential trend in terms of spread of COVID-19. 

**What are some patterns that can signal a major upcoming wave? (e.g. if we see a steady growth in cases/deaths for x consecutive days, does that tell us a bigger picture about the state of COVID-19 in the US or a specific state?)**

We found certain patterns within the data of rise and falls, but could not conclusively discern a reason for the trends. Some of the trends were due to data reporting issues and when we smoothed out for the trends by taking the 7-day average, some of those disappeared. Again, we initially hypothesized that major events in the United States led to increases to COVID-19, although true in its locality, does not detract from the exponential growth we are seeing today.

## Additional Insight

In our work we also tried to attribute local policy changes in regards to COVID-19. Laws were extremely strict initially in March. As the hysteria died, people across the US began to lax their weariness in regards to danger and spread of COVID-19. This can also be reflected in local laws. Phased approaches of openings were announced and laws began to relax across the board. Potentially, the initial fear and stronger laws could have led to the steadiness of the summer spread. However, the relaxed laws and the normalization of COVID-19 in our lives have led to a surge in cases, which we see today. Below is an example of NY law changes overtime. The laxing of restrictions in the summer months are clearly reflected in the timeline (https://en.wikipedia.org/wiki/COVID-19_pandemic_in_New_York_(state)):


```{r}
library(rvest)
library(knitr)
url <- 'https://en.wikipedia.org/wiki/COVID-19_pandemic_in_New_York_(state)'
webpage <- read_html(url)
tbls <- html_nodes(webpage, "table")
df <- html_table(tbls[grep("Timeline for",tbls,ignore.case = T)],fill = T)[[1]]
kable(df, format = "pipe")
```

## Future Work and Potential Developments

Although our project is depicting a gloomy outlook for the future of the US, we have identified future developments that can potentially stem the tide and control the the spread of COVID-19. [Pfizer has been approved for distribution](https://www.fda.gov/emergency-preparedness-and-response/coronavirus-disease-2019-covid-19/pfizer-biontech-covid-19-vaccine) of the COVID-19 vaccine and expects to have [50 million doses](https://www.pfizer.com/products/coronavirus/manufacturing-and-distribution) available by the end of 2020. This is a great news and can potentially influence the trend. At the same time, we understand that widespread distribution and adoption are still slow processes. With this in mind, the exponential trend will still continue for the foreseeable future. Once the vaccine becomes more accessible to the everyday person,  we can then monitor the trends and see how effective the vaccine are. As is the case with new drugs, the effectiveness is still yet to be determined. 

## Final Words

In concluding our project, we hope that you gained a better understanding of the spread of COVID-19 and urgency of the situation. Given the exponential growth in the recent months, we can see the effects of our behaviors and our local laws have consequences months later. We hope that you came away more informed about the situation.
